MSc Thesis

### 2023-03-21 manuscript/main.tex
worked on discussion

\documentclass[12pt, letterpaper, ]{article}
\usepackage[style=authoryear, maxnames=2, maxbibnames=99]{biblatex}
\addbibresource{merged_biblio2.bib}
\addbibresource{internet_stuff.bib}
%\addbibresource{induction.bib}
%\addbibresource{gwas_silicon.bib}
\usepackage{caption}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{url}
\usepackage{lineno}
\linenumbers
\graphicspath{./images/}
\geometry{
        letterpaper,
        top = 1.27cm,
        bottom = 1.27cm,
        left = 2.54cm,
        right = 2.54cm,
        }
\title{MSc Thesis}
\author{Isaac Peetoom Heida}
\date{December 2022}
\doublespace

@@

To test the prevalence of rapid silicon accumulation in canadian cereal crops, I selected three cultivars for each of oats, bread wheat, triticale, and barley. I selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations. At the start of the experiment, I germinated seeds in germination trays filled with moist sand. After four days, I transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid. Though potting mix and fresh water contain some amount of plant available silicon, I added the silicic acid to ensure that there would be no silicon limitation to the plants. I randomized the location of each pot within the growing space. A flood table bottom watered the pots with nutrient solution. I assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory. I simulated herbivory by application of 1 mM MeJA solution to the entire above-ground portion of the plant (Waterman et al. 2021b), while crickets housed in water-pik tubes provided true herbivory. Prior to introduction to the the plants, I acclimated crickets by feeding them on the same species used in this trial. Immediately preceding cricket application, I placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.

@@

18 hours after treatment application, I harvested leaf material by clipping all undamaged (i.e. no cricket exposure) leaves. I harvested three fully expanded upper leaves the plants, and split the plants in half along the midvein of the leaf. I placed one half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it. I plaed the other half of the tissue into a coin envelope, oven dried it for 4 days at 60$\deg$ C, transferred it to a 2ml microcentrifuge tube with three 3.2mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz). After freeze drying, the samples for phenolic processing were ground under the same conditions.

@@

To test all three of our hypotheses, I used linear mixed effects models implemented in (\cite{douglas_bates_2023_7529836}) in Julia 1.8.5 (\cite{bezanson2017julia}). Prior to running our full model, I first tested for an effect of biomass of silicon content, as in other defensive pathways plant size shows corelations with defense levels \cite{carmona_plant_2011}. We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model. Thus, we tested the effects of our species and induction treatments on measured leaf silicon using the following model formula:

@@

\subsection{Results}
Oats had the lowest silicon content at 0.34\%. Overall, methyl-jasmonate and insect treatments increased plant silicon by [x] and [y]\% respectively. Our model showed that both species and treatment type had effects on the plant silicon content. Parameter estimates and 90\% credible intervals are summarized in Table \ref{Tab:params}. Species and induction treatments had an interaction.

@@

@@

\begin{figure}[h]
        \includegraphics[width = \textwidth]{images/induction_plot.png}
        \centering
        \caption{The effects of crop species and induction treatment on leaf silicon content. Plants were treated either with a 1mM methyl jasmonate spray, or exposure to house crickets \textit{Acheta domesticus}. Leaves were sampled 24 hours after treatment, and were analyzed using XRF.}
\end{figure}

>>>>

\documentclass[12pt, letterpaper, ]{article}
\usepackage[style=authoryear, maxnames=2, maxbibnames=99]{biblatex}
\addbibresource{merged_biblio2.bib}
\addbibresource{internet_stuff.bib}
%\addbibresource{induction.bib}
%\addbibresource{gwas_silicon.bib}
\usepackage{caption}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{url}
\usepackage{lineno}
\usepackage{gensymb}
\linenumbers
\graphicspath{./images/}
\geometry{
        letterpaper,
        top = 1.27cm,
        bottom = 1.27cm,
        left = 2.54cm,
        right = 2.54cm,
        }
\title{MSc Thesis}
\author{Isaac Peetoom Heida}
\date{December 2022}
\doublespace

@@

To test the prevalence of rapid silicon accumulation in canadian cereal crops, I selected three cultivars for each of oats, bread wheat, triticale, and barley. I selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations. At the start of the experiment, I germinated seeds in germination trays filled with moist sand. After four days, I transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid. Though potting mix and fresh water contain some amount of plant available silicon, I added the silicic acid to ensure that there would be no silicon limitation to the plants. I randomized the location of each pot within the growing space. A flood table bottom watered the pots with nutrient solution. I assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory. I simulated herbivory by application of 1 mM methyl-jasmonate solution to the entire above-ground portion of the plant (Waterman et al. 2021b), while crickets housed in water-pik tubes provided true herbivory. Prior to introduction to the the plants, I acclimated crickets by feeding them on the same species used in this trial. Immediately preceding cricket application, I placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants. Prior to harvest, I recorded whether the crickets had initiated feeding on the plants by visual inspection of the leaves for missing tissue.

@@

18 hours after treatment application, I harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein. I placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for XRF analysis. I placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it. After freeze drying, the samples for phenolic processing were ground under the same conditions as the other samples.

@@

Despite the starvation, some crickets did not initiate feeding during the exposure period. I filtered out plants assigned to the insect induction treatment that recieved no damage, to ensure that they would not confound the model. Prior to running our full model, I first tested for an effect of biomass of silicon content, as other defensive pathways show correlations between plant size and defense levels \cite{carmona_plant_2011}. We found a negative correlation between plant size and silicon content ($\beta = -0.067 \pm 0.017, p < 0.001$), and thus included plant size as a covariate in our final model. To test all three of our hypotheses, I used linear mixed effects models and tested the effects of our species and induction treatments on measured leaf silicon using the following model formula:

@@

\subsubsection{Software used}

I compiled the final dataset using \verb|DataFrames.jl| (\cite{bogumil_kaminski_2023_7632427}) in Julia 1.8.5 (\cite{bezanson2017julia}). I implemented the biomass regression using \verb|GLM.jl| (\cite{douglas_bates_2023_7529836}). I tested the full mixed effects model in R 4.2.2 (\cite{r_core_team_2022}) using \verb|lme4| (\cite{lme4_bates_2015}), and performed a post-hoc tukey test using \verb|emmeans| (\cite{lenth_2023_emmeans}). I generated graphics in Julia using \verb|Plots.jl| (\cite{tom_breloff_2023_7736124}).

\subsection{Results}
Among the various cultivars, average uninduced silicon content ranged from 0.26\% to 0.91\% (Figure \ref{Fig:baseline_si}). Amongst species, oats had the lowest average silicon content at $0.34 \pm 0.02\% (\mu \pm SE)$, while wheat had the highest average silicon content at $0.76 \pm 0.05\% (\mu \pm SE)$ (Figure \ref{Fig:baseline_si}). Counter to my predictions, I failed to find strong support for inducible increases in silicon content among the tested plant species. Despite a small p-value in the ANOVA table of the model ($p = 0.046, F = 3.14, df = 2,185.9$), the model showed only moderate support for an effect of methyl-jasmonate application ($\beta = 0.079 \pm 0.043, p = 0.070$), and no support for an effect of cricket exposure on silicon content ($\beta = 0.038 \pm 0.050, p = 0.453$). I found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,185.9$). Post-hoc tukey tests revealed that this was driven primarily by wheat's response to my induction treatments, where both induction treatments were associated with decreased silicon content (methyl-jasmonate: $p = 0.052, t = -1.95, df = 186$, Cricket: $p = 0.00184, t = -3.16, df = 186$) (Figure \ref{Fig:induction}). %Our model showed that both species and treatment type had effects on the plant silicon content. Parameter estimates and 90\% credible intervals are summarized in Table \ref{Tab:params}. Species and induction treatments had an interaction.

@@

Recent research in inducible silicon plant defenses has focussed on the short-term dynamics of silicon uptake (\cite{waterman_short-term_2021,waterman_short-term_2021-1}). The promising results of this work has been highlighted for its potential applications in agriculture, where sensitive and rapid defensive phenotypes could improve plant performance and reduce reliance on more intensive pest-control measures. In this study, I attempted to demonstrate rapid silicifcation in four cereal crops. I failed to find evidence of rapidly induced silicon uptake in response to either methly-jasmonate application or herbivore exposure. In our study, the variation among our cultivars within species was similar in magnitude to variation among species (Figure \ref{Fig:baseline_si}), possibly obscuring the effects of our treatments. This study also differed from previous studies demonstrating rapid silicon uptake in a number of ways. The two previous studies both grew plants in liquid nutrient solution, carefully standardized to maintain consistently high silicon availability. In my study, I grew plants in potting soil amended with solid silicic acid. In natural soil environments, the majority of plant-available silicon is derived from mineral or biogenic sources, and thus requires dissolution into the soil solution. The theoretical maximum concentration of silicic acid in soil solution is 2 mM, however observed concentrations can be much below that. In my growing conditions, I applied silicic acid in excess of the average availability of phytoliths (a major source of plant-avaialbe silicon), so as to avoid soil conditions with low silicon presense. Despite this, silicon availability in the soil solution, and dissolution rates from solid to aqueous forms, may have been too low to meet the demands of rapid silicon uptake.

@@

\begin{figure}[h]
        \includegraphics[width = \textwidth]{images/spp_si_content.png}
        \centering
        \caption{Baseline (uninduced) silicon content in the cereal cultivars used in this study. Cultivar species is notated in parentheses in the legend.}
        \label{Fig:baseline_si}
\end{figure}

\begin{figure}[h]
        \includegraphics[width = \textwidth]{images/induction_plot.png}
        \centering
        \caption{The effects of crop species and induction treatment on leaf silicon content. Plants were treated either with a 1mM methyl jasmonate spray, or exposure to house crickets \textit{Acheta domesticus}. Leaves were sampled 24 hours after treatment, and were analyzed using XRF.}
        \label{Fig:induction}
\end{figure}
